rethinking and the 5 genera

### 2021-03-16 scripts/56_rethinking_genera.Rmd
feeling I didn something major cuz never did this before pmap within pmap

bh <- meta %>% 
  arrange(-Bacteroides) %>% 
  slice(1) %>% 
  select(Akkermansia:Ruminococcus)
```

```{r}
pl <- inv_logit( post$ba * (-1.612806	) + post$bb * (-5.69897		) + post$be * (-5.69897	) +  post$bf * (-5.69897		) + post$br * (-0.5244935) + post$a[, 1] )

@@

bind_rows(
  tibble(
    value = pl,
    grp = 'low'
  ),
  tibble( 
    value = ph,
    grp = 'high'
  )
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')

@@

bind_rows(
  tibble(
    value = pl,
    grp = 'low'
  ),
  tibble( 
    value = ph,
    grp = 'high'
  )
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')

@@

>>>>

bh <- meta %>% 
  arrange(-Bacteroides) %>% 
  slice(1) %>% 
  select(Akkermansia:Ruminococcus)

pl <- inv_logit( post$ba * (-1.612806	) + post$bb * (-5.69897		) + post$be * (-5.69897	) +  post$bf * (-5.69897		) + post$br * (-0.5244935) + post$a[, 1] )

@@

bind_rows(
  tibble(
    value = pl,
    grp = 'low'
  ),
  tibble( 
    value = ph,
    grp = 'high'
  )
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
              xlab = 'Predicted probability of toxicity', ylab = 'Probability density')

@@

bind_rows(
  tibble(value = pl,grp = 'low'),
  tibble( value = ph,grp = 'high')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
              xlab = 'Predicted probability of CR d100', ylab = 'Probability density')

@@

```{r}
# ruminococcus relab quantile
meta_ <- meta %>% 
  select(Akkermansia:Ruminococcus)

# the post coeffs in df format
postdf <- bind_cols(ba = post$ba, bb = post$bb, be = post$be, bf = post$bf, br = post$br, am = post$a[,1])

# to responde to Sean's comments to not use max and min "extreme"patient data
# use the 90% and 10% quantile patient data
# Jonas: how do you know that the patient falling onto the 90%tile is not “exceptional” and therefore an “outlier” in another feature that your model includes, besides ruminococcus, he suggests to sample in that 0-10% quantile range patient data for example
# It is 100 jointly sampled coefficients multiplied with the composition of bacteria in the current patient
N <- 100
# for the top and bottom quantile range of ruminococcus relab
ru_top <- meta_ %>% 
  filter(Ruminococcus >= quantile(meta$Ruminococcus, 0.9))
ru_bot <- meta_ %>% 
  filter(Ruminococcus <= quantile(meta$Ruminococcus, 0.1))

# sample in the top and bot patients 100 times
ru_top_samp <- ru_top %>% 
    sample_n(size = N, replace = T)
ru_bot_samp <- ru_bot %>% 
    sample_n(size = N, replace = T)

# a function for every patient data generate 100 post prob of CR (sampling from the post 100 times)

per_pt_sample_post_prob <- function(lga_, lgb_, lge_, lgf_ , lgr_){
  # the input is the log relab of the 5 genera for that patient
  post_samp <- postdf %>% 
    sample_n(size = N, replace = F)

  ret = pmap(post_samp, function(ba, bb, be, bf, br, am) {
    inv_logit( ba * lga_ + bb * lgb_ + be * lge_ +  bf * lgf_ + br * lgr_ + am ) 
  }) %>% 
    set_names(seq(1, N)) %>% 
    bind_rows() %>% 
    gather()
  return(ret)
}

# for the top quantile patients 
res_top <- pmap(ru_top_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
}) %>% 
  set_names(paste('P', seq(1, N), sep = '')) %>% 
  bind_rows(.id = 'ptsamp')

# for the bottom quantile patients
res_bot <- pmap(ru_bot_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
}) %>% 
  set_names(paste('P', seq(1, N), sep = '')) %>% 
  bind_rows(.id = 'ptsamp')
```

```{r}
# plotting and visualizing
res_top %>% 
  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
              title = '100 sampled patients that had top 10% Ruminococcus relab',
              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
  theme(legend.position = 'none')
```

```{r}
res_bot %>% 
  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
              title = '100 sampled patients that had bottom 10% Ruminococcus relab',
              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
  theme(legend.position = 'none')
```
